---
Title: R Skills Assessment
author: "MEES 614"
date: "Monday, February 2, 2015"
output: html_document
---

## Student UID: 113145995

# 1. Making objects, assigning values, and simple arithmetic

Make three objects named *x, y, and x*.  Assign any numerical value to each & then multiply them together.
```{r}
x<-c(8)
y<-c(10)
z<-c(3)

v<-x*z*y
v
```

# 2. Creating vectors

Create three vectors. Name the first *myVec* and have it contain any 5 numbers. Name the second *myRep* and have it contain the number 10 repeated 217 times. The third should be named *mySeq* and contain a sequence of numbers from 1-100, counting by 0.25. Have R return the length of *mySeq*.
```{r}
myVec<-c(1,2,3,4,5)
myRep<-rep(10, 217)
mySeq<-seq( from=1, to=100, by=0.25 )
length(unique(mySeq))
```

# 3. Working with objects, files, directories 
 
Execute functions to:

* list the objects in your workspace
* print your current working directory
* list the files and folders in your current working directory
* save your workspace

```{r}
ls()
getwd()
#save.workspace()
```

# 4. Loading, saving, plotting, and indexing data 

Load the iris dataset & do the following:

  * Make a new object that has the value in row 10, column 1.
  * Assign new values to the 5th row of the iris dataset by multiplying the current values by 2.
  * Make a new object containing rows 1-10, 15, 35-40 and columns 1-2, 5 from the iris dataset.
  * Make a new object that contains everything except the 10th row of the iris dataset.
  * Make a new object that contains only those records with a sepal length greater than 5.
  * Make a new object by subsetting the iris dataset so that it contains only the records for **setosa** and save this object to your hard drive as a csv. 
  * Plot *Sepal Length* vs. *Sepal Width* from the full iris dataset.

```{r}
##1
data(iris)
t<-iris[10,1]
str(iris)
#2?
s<-iris[5, ]
s2<-s*2
#cannot do this with species (=factor)
#3?
i1<-subset(iris[1:10,1:2])
i2<-subset(iris[15,1:2])
i3<-subset(iris[35:40:10,1:2])
#i4<-cbind(i1,i2,i3)

#4
iris2<-iris
iris2<-iris[-10, ]
#5
newobject2<-iris$Sepal.Length>5
b<-iris$Sepal.Length[ iris$Sepal.Length > 55 ]
#data<-cbind(iris$Sepal.Length, iris$Sepal.Width)

#6 
iris3 <- subset(iris, iris$Species=="setosa")
#7
plot(iris$Sepal.Length~iris$Sepal.Width )




```

# 5. Lists and Dataframes 

Create a list that contains *myVec*, *myRep*, and *mySeq*. Create a dataframe from the first five values of *myVec*, *myRep*, and *mySeq*. Have R return the dimensions and structure of the dataframe and, separately, the number of rows, columns, and column names. Save the dataframe as a csv table.  
```{r}
#data.frame(cbind(myVec, myRep, mySeq)
           
```

# 6. Installing and Loading packages

Install and then load the *dismo* package.
```{r}
#library(dismo)

```

# 7. Plotting 

A. Create a histogram of the sepal length data from the iris dataset and add to it:

* A density curve
* A normal curve
* A legend

B. Load and plot the lynx dataset, then create a new plot as follows:

*Make a combined point-line plot, using character code 21 for the points and a line width of 3 for the segments.
*Make the line segments and outline of the points blue.  Make the fill of the points grey.
*Label the y-axis *Count* and the x-axis *Year*
*Lastly, plot the new plot and the original side-by-side in the same window.

```{r}
hist(iris$Sepal.Length)
```

# 8. Writing for-loops

A. Write a for-loop the prints "1. Hello world!" five times, replacing the numer with the current interation of the for-loop.

B. Write a nested for-loop that assigns a random number drawn from a uniform distribution (see ?runif) between 0-1 to each position within a 25 x 25 matrix named *myMat*. Use the *image* function to plot a heatmap of your matrix. Create an object that has only the values in your matrix that are greater than 0.5.

```{r}
#M <- matrix( data=NA, nrow=5, ncol=1 )

#for(i in 5:1){M[ ,i]<- c("Hello world!")} M


```

# 9. Using *apply* functions 

Use the apply function to calculate the row means of *myMat* from #8 above.
```{r}
#ddply(.data=myMat, .variables="8", .fun="mean"")

```

# 10. Writing custom functions

Write a simple function that will calculate the cube of any number. Plot the results for values 0-10.

```{r}
# simple function to return square of a number

```

# 11. If-else statements

Write an if-else statement to calculate the price of a speeding ticket, with the following rules:

1. The ticket fee is $25 for every mile per hour above 55.
2. Plus an additional $50 for every mile per hour above 65. 
3. There is an additional $100 charge for red cars.

*What is the ticket fee for a red car traveling 75 mph?
*What is the ticket fee for a black car traveling 58?

```{r}

```

# 12. Finishing up

Save your workspace & then clear it.

```{r}

```



